Active Learning from Blackbox to Timed Connectors

### 2016-01-16 paper/paper.tex
.

\usepackage{amssymb}
\usepackage{amsmath}
\usepackage{cite}
\usepackage[normalem]{ulem}
% for a smile-face character
\usepackage{wasysym}
% listing for codes
\usepackage{listings}
% listing for golang
% NOTE listings-golang.sty is required
\usepackage{listings-golang}
% to support flow graphs
\usepackage{textcomp}
% highlight in tables
\usepackage{xcolor}
\usepackage{colortbl}

\usepackage[linesnumbered, ruled, vlined]{algorithm2e}

@@

\begin{algorithm} 
  \caption{BuildTable} 
  \label{alg:buildtable}
  \KwIn{Oracle interface $mq$, Input actions $\mathcal{I}$, suffix set $D$}
  \KwOut{Observation table $obs$} 
  $obs$ initialized as empty\;
  $unclosed=\{\varepsilon\}$\;
  \Repeat{$unclosed=\varnothing$}
  {
    $next=\{st|\forall s\in unclosed,\forall t\in \mathcal{I}\}$\;
    append $unclosed$ to $obs$\;
    $unclosed=\{seq\in next|\forall a\in obs, seq \not\sim_D a\}$\;
  }
  \Return $obs$\; 
\end{algorithm}

@@

\begin{figure}[ht]
  \begin{center}
    \begin{displaymath}
      \begin{array}{l||ccccc}
        \hline
        & a,\rread & \bot,\rread & a,\rnoread & \bot,\rnoread & T\\
        \hline\hline
        \varepsilon & \rblock & \rblock & \rempty & \rempty & \rempty \\
        \hline
        a,\rread & \rblock & \rblock & \rempty & \rempty & \rempty \\
        \bot,\rread & \rblock & \rblock & \rempty & \rempty & \rempty \\
        \rowcolor[gray]{.9}
        a,\rnoread & \rblock & \rblock & \rblock & \rempty & \rempty \\
        \bot,\rnoread & \rblock & \rblock & \rempty & \rempty & \rempty  \\
        T & \rblock & \rblock & \rempty & \rempty & \rempty \\
        \hline
      \end{array}
    \end{displaymath}
    \input{./images/hypo.tex}
  \end{center}
  \caption{Observation Table and Corresponding Hypothesis}
  \label{fig:hypo}
\end{figure}

@@

\begin{figure}[ht]
  \begin{center}
    \begin{displaymath}
      \begin{array}{l||ccccc}
        \hline
        & a,\rread & \bot,\rread & a,\rnoread & \bot,\rnoread & T\\
        \hline\hline
        \varepsilon & \rblock & \rblock & \rempty & \rempty & \rempty \\
        \hline
        a,\rread & \rblock & \rblock & \rempty & \rempty & \rempty \\
        \bot,\rread & \rblock & \rblock & \rempty & \rempty & \rempty \\
        a,\rnoread & \rblock & \rblock & \rblock & \rempty & \rempty \\
        \bot,\rnoread & \rblock & \rblock & \rempty & \rempty & \rempty  \\
        T & \rblock & \rblock & \rempty & \rempty & \rempty \\
        \hline
        a,\rnoread-a,\rread & \rblock & \rblock & \rblock & \rempty & \rempty \\
        a,\rnoread-\bot,\rread & \rblock & \rblock & \rblock & \rempty & \rempty \\
        a,\rnoread-a,\rnoread & \rblock & \rblock & \rblock & \rempty & \rempty \\
        a,\rnoread-\bot,\rnoread & \rblock & \rblock & \rblock & \rempty & \rempty \\
        a,\rnoread-T & \rblock & \rblock & \rblock & \rempty & \rempty \\
        \hline
      \end{array}
    \end{displaymath}
  \end{center}
  \caption{A Closed Observation Table}
  \label{fig:hypo2}
\end{figure}

\subsection{Counter-Examples' Analysis} 
Apparently, the closed hypothesis presented in Figure \ref{fig:hypo2} is not equal to the 2-Timer
channel. It's easy to find a counter-example $s=a,\rnoread-T-T-\bot,\rread$ where $mq(s)=a$ while
according to the hypothesis, the result shoud have been $\rempty$. In this section, we show how to
find and analyze counter-examples, which is mainly based on \cite{DBLP:conf/sfm/SteffenHM11}.

@@

Considering a counter example $s$, we have
\[
\mbox{MaxSuffix}(s)=\max_d\{s=s'd, mq(acc(s')d)\neq mq(s)\}
\]
Since $mq(acc(s))=hq(s)\neq mq(s)$ and $mq(s) = mq(s)$, the existance of such $d$ is easy to prove.
After appending $d$ in $D$, we explain the closedness will be violated.

The whole L* algorithm can be concluded as follows.
\begin{algorithm} 
  \caption{L*} 
  \label{alg:lstar}
  \KwIn{Oracle interface $mq,eq$, Input actions $\mathcal{I}$}
  \KwOut{Observation table $obs$} 
  $D=I$\;
  \Repeat{$ce=true$}{
    $obs$ = BuildTable($mq$, $\mathcal{I}$, $D$)\;
    $ce=eq(obs)$\;
    \If {$ce!=true$} {
      D.append(MaxSuffix(s))\;
    }
  }
  \Return $obs$\; 
\end{algorithm}

@@

Both \emph{Reo Coordination Models} and \emph{Adapted L* Algorithm} are implemented in
Golang\cite{golang}.

Golang (or Google Go) is a rising programming language started by Google Inc. The language is widely
known for its elegant design and impressive efficiency. Moreover, the concurrency model of Golang
comes from CSP\cite{DBLP:books/ph/Hoare85}. As a channel-based model, CSP shares a similar idea with
Reo and makes our implementation much more natural.

@@

As shown in Table \ref{tabel:timecost}, time consumption mainly comes from membership queries.
Since time is involved in our model, it's inevitably that simulation takes time to behave normally.
Even worse, since our models are treated as blackbox. With no access to inner behaviour of the
connectors, it's almost impossible to accelerate the simulation process.

Fortunately, there are still other optimization solutions. After reviewing our algorithm, we found
that simulations on similar sequences were invoked frequently:

>>>>

\usepackage{amssymb}
\usepackage{amsmath}
\usepackage{cite}
\usepackage[normalem]{ulem}
% for a smile-face character
\usepackage{wasysym}
% listing for codes
\usepackage{listings}
% listing for golang
% NOTE listings-golang.sty is required
\usepackage{listings-golang}
% to support flow graphs
\usepackage{textcomp}
% highlight in tables
\usepackage{xcolor}
\usepackage{colortbl}
% used to encode algorithms
\usepackage[linesnumbered, ruled, vlined]{algorithm2e}
\usepackage{array}

@@

\begin{algorithm} 
  \caption{CloseTable} 
  \label{alg:buildtable}
  \KwIn{Oracle interface $mq$, Input actions $\mathcal{I}$, suffix set $D$}
  \KwOut{Observation table $obs$} 
  $obs$ initialized as empty\;
  $unclosed=\{\varepsilon\}$\;
  \Repeat{$unclosed=\varnothing$}
  {
    $next=\{st|\forall s\in unclosed,\forall t\in \mathcal{I}\}$\;
    append $unclosed$ to $obs$\;
    $unclosed=\{seq\in next|\forall a\in obs, seq \not\sim_D a\}$\;
  }
  \Return $obs$\; 
\end{algorithm}

@@

\begin{figure}[ht]
  \begin{center}
    \begin{displaymath}
      \begin{array}{l||ccccc||c}
        \hline
        & a,\rread & \bot,\rread & a,\rnoread & \bot,\rnoread & T & \mbox{AccessSequence}\\
        \hline\hline
        \varepsilon & \rblock & \rblock & \rempty & \rempty & \rempty & \varepsilon\\
        \hline
        a,\rread & \rblock & \rblock & \rempty & \rempty & \rempty & \varepsilon\\
        \bot,\rread & \rblock & \rblock & \rempty & \rempty & \rempty & \varepsilon \\
        \rowcolor[gray]{.9}
        a,\rnoread & \rblock & \rblock & \rblock & \rempty & \rempty & \mbox{NO MATCH}\\
        \bot,\rnoread & \rblock & \rblock & \rempty & \rempty & \rempty & \varepsilon \\
        T & \rblock & \rblock & \rempty & \rempty & \rempty \varepsilon & \varepsilon \\
        \hline
      \end{array}
    \end{displaymath}
    % \input{./images/hypo.tex}
  \end{center}
  \caption{Observation Table and Corresponding Hypothesis}
  \label{fig:hypo}
\end{figure}

@@

\begin{figure}[ht]
  \begin{center}
    \begin{displaymath}
      \small
      \begin{array}{l||ccccc||c}
        \hline
        & a,\rread & \bot,\rread & a,\rnoread & \bot,\rnoread & T & \mbox{AccessSequence}\\
        \hline\hline
        \varepsilon & \rblock & \rblock & \rempty & \rempty & \rempty & \varepsilon \\
        \hline
        a,\rread & \rblock & \rblock & \rempty & \rempty & \rempty & \varepsilon\\
        \bot,\rread & \rblock & \rblock & \rempty & \rempty & \rempty & \varepsilon \\
        a,\rnoread & \rblock & \rblock & \rblock & \rempty & \rempty & a,\rnoread \\
        \bot,\rnoread & \rblock & \rblock & \rempty & \rempty & \rempty & \varepsilon \\
        T & \rblock & \rblock & \rempty & \rempty & \rempty & \varepsilon \\
        \hline
        a,\rnoread-a,\rread & \rblock & \rblock & \rblock & \rempty & \rempty & a,\rnoread \\
        a,\rnoread-\bot,\rread & \rblock & \rblock & \rblock & \rempty & \rempty & a,\rnoread \\
        a,\rnoread-a,\rnoread & \rblock & \rblock & \rblock & \rempty & \rempty & a,\rnoread \\
        a,\rnoread-\bot,\rnoread & \rblock & \rblock & \rblock & \rempty & \rempty & a,\rnoread \\
        a,\rnoread-T & \rblock & \rblock & \rblock & \rempty & \rempty & a,\rnoread \\
        \hline
      \end{array}
    \end{displaymath}
  \end{center}
  \caption{A Closed Observation Table}
  \label{fig:hypo2}
\end{figure}

\subsection{Counter-Examples' Analysis} 
Apparently, the closed hypothesis presented in Figure \ref{fig:hypo2} is not equal to the 2-Timer
channel. It's easy to find a counter-example $s=a,\rnoread-T-T-\bot,\rread$ where $mq(s)=a$ while
according to the hypothesis, the result shoud have been $\rempty$. In this section, we show how to
find and analyze counter-examples using the method in \cite{DBLP:conf/sfm/SteffenHM11}.

@@

Considering a counter example $s$, we have
\[
\mbox{MaxSuffix}(s)=\max_d\{s=s'd, mq(acc(s')d)\neq mq(s)\}
\]
Since $mq(acc(s))=hq(s)\neq mq(s)$ and $mq(s) = mq(s)$, the existance of such $d$ is easy to prove.
After appending $d$ in $D$, we can rebuild a closed hypothesis where $s$ is no longer a
counter-example.

The whole L* algorithm can be concluded as follows.
\begin{algorithm} 
  \caption{L*} 
  \label{alg:lstar}
  \KwIn{Oracle interface $mq,eq$, Input actions $\mathcal{I}$}
  \KwOut{Observation table $obs$} 
  $D=I$\;
  \Repeat{$ce=true$}{
    $obs$ = CloseTable($mq$, $\mathcal{I}$, $D$)\;
    $ce=eq(obs)$\;
    \If {$ce!=true$} {
      D.append(MaxSuffix(s))\;
    }
  }
  \Return $obs$\; 
\end{algorithm}

@@

Both \emph{Reo Coordination Models} and \emph{Adapted L* Algorithm} are implemented in
Golang\cite{golang}, which is widely known for its elegant design and impressive efficiency.
Moreover, with a CSP-style\cite{DBLP:books/ph/Hoare85} cocurrency model,\texttt{Golang} shares many
similar ideas with Reo, and in turn makes our implementation much more natural.

@@

As shown in Table \ref{tabel:timecost}, time consumption mainly comes from membership queries.
With time involved, every single membership query takes a lot of time inevitably.
After reviewing our algorithm, we found that simulations on similar sequences were invoked frequently:
